Convergence Studies on Mass Transport in Disks with Gravitational Instabilities. \\ I. The Constant Cooling Case

### 2010-12-22 Paper_1.tex
Scott's first round of revisions.
	modified:   Paper_1.tex
	modified:   general.bib

We conducted a convergence study of the azimuthal resolution in 3D hydrodynamic simulations of a protostellar disk susceptible to gravitational instabilities (GIs). 
The disk was primarily studied during its asymptotic state, when
heating and cooling are in balance. The cooling was treated as a constant global cooling time.
Four simulations were conducted, identical except for the number of azimuthal grid 
points, $l_{max} = $ 64, 128, 256 and 512, used in the calculations. 
The strengths of the azimuthal modes, gravitational torques, and the effective \citet{shakura1973} $\alpha$ 
arising from gravitational stresses were determined for each resolution.  We examined the resolution-dependence of each quantity
and evaluated both the converged modal amplitude distribution and the effective value of $\alpha$
as $l_{max} \rightarrow \infty$.   Non-axisymmetric amplitude was found in all modes and 
the amplitude summed over all modes was essentially independent of resolution.
The modal distribution of amplitudes, however, was significantly affected by resolution.  
The size of the azimuthal grid, $l_{max}$,
limits the number of modes that can be accurately measured to $l_{max}/2$.  Higher resolution simulations therefore
have more modes available to them.  Non-axisymmetric structure is present in all available modes.  As a consequence,
higher resolution simulations have a larger fraction of their total amplitude in high-order modes.
High-order modes
are more confined to act on local structure than low-order modes, therefore as the resolution increases and low-order amplitudes decrease the total gravitational stresses decreases as well.
As a result, higher resolution simulations experience 
weaker gravitational torques than lower resolution simulations.   The effective  $\alpha$
also depends upon the magnitude of the stresses.  Thus
$\alpha_{eff}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
The converged $\alpha_{eff}$ is consistent with
predictions from an analytic local theory for thin disks by \citet{gammie2001}, 
but only over many dynamic times when averaged over a substantial volume of the disk.
Results obtained in the $l_{max} = 512$ simulations differ with converged values
by less than 20\%.

@@

Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
modes can become unstable in the linear regime \citep{nelson1998, pickett1998, lodato2004,boley2006,cossins2009}.  Even if the initial growth is dominated by one mode, numerous modes usually appear in the nonlinear regime. As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, referred to as the {\it asymptotic state},
then power will exist at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.  
In the asymptotic state, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
\citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.  Here convergence studies are important because 
higher order modes dissipate energy on shorter wavelengths than lower order modes.  The mix of low and high order modes available to a disk is set by the resolution of the simulation. {\bf Multiple issues with this paragraph. Didn't want to rewrite unilaterally. 1) In the first couple sentences the description of mode growth needs some tweaking, many modes appear in the linear regime, and although we have seen amplitude at all m values I don't know that it's a given. 2) I wouldn't describe the structures formed by GI activity as ``strongly nonlinear turbulence''. 3) In reviewing Gammie 2001 it seems that he is actually referring to a turbulent cascade (a fact I hadn't fully appreciated until now) I think it's pretty clear that the effective alpha we see is, in fact, not due to a turbulent cascade. i.e. alpha is driven by low order coherent modes. We should discuss this at some point}

\citet{shakura1973} proposed turbulence as the primary source of effective viscosity in gas disks
and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  The source of the 
turbulence in their $\alpha$-disk formulation is unspecified. Although their original work was directed towards X-ray bright
accretion disks around black holes, this heuristic approach has proven fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling. In the 
$\alpha$-disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $\Sigma$ is the surface density and $H$ represents
the disk scale height. The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
and heating are dominated by local processes}. Lin and Pringle (1987, 1990) were among the first to suggest
that GI-induced transport could be described within a viscous
framework.  However, the  validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation. 
The locality of dissipation in GI-active disks has not been rigorously tested and remains an open question.

Several studies support the idea that angular momentum transport by GIs in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.  This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance, i.e., disks subject to gravitoturbulence,
 it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.  In particular, 
an $\alpha$-viscosity may be consistent
with transport in GI acitve disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
in systems with larger $M_d/M_*$ are dominated by lower-order modes that act in a more global sense and thus cannot be well represented by an $\alpha$-approach
\citep{lodato2004, cossins2009, vorobyov2010}.

\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating caused by GIs was balanced by the local cooling rate.  
Assuming a razor thin gravitationally unstable disk where the heating caused by
GIs was balanced by a local cooling rate of $t_{cool}\Omega$.
Gammie found
$\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}$,
where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
The two dimensional index relates
to the three dimensional index, $\Gamma$, by 
$\gamma = (3\Gamma - 1) / (\Gamma + 1)$
for a non-self-gravitating
disk and
$\gamma = 3 - 2/\Gamma $
for a strongly self gravitating disk.   \cite{lodato2004} derived a similar result when considering three dimensional protoplanetary disks.
Motivated by
the results from \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order, \citet{kratter2008} 
proposed $\alpha_{eff} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
are functions of $Q$ and $M_d/\left( M_d+M_*\right)$ representing 
the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.

We address these, and similar questions, using a grid-based, finite-difference 3D hydrodynamics code
with self gravity.
For computational grids of a fixed {\it physical} size,
grids with more elements provide finer meshes, and thus a more accurate solution. 
However, increased accuracy is offset by computational penalties.
The number of equations solved at each timestep 
increases as the grid size is increased.  At the same time, the Courant condition 
forces shorter timesteps.  A mesh convergence study 
provides a common approach to balancing accuracy and computing resources.
In the following sections we follow the evolution of a disk subject to GIs
using four different
resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
may be particularly important, because higher azimuthal resolution 
allows non-axisymmetric amplitude to spread to higher-order modes 
that behave more locally. We test this by examining the amplitude in various 
modes to see how they are affected by the choice of grid.
In the process, 
we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium If this is the case, we investigate how closely these disks approximate the gravitoturbulence of \citet{gammie2001} and examine the distribution of $\alpha_{eff}$. To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.  Simulations are performed using four different resolutions in the azimuthal direction.

@@

The initially unstable disk model passed through several phases of evolution \citep[see also][]{pickett2003, mejia2005}.  The initial disk is unstable to the growth of non-axisymmetric structures. During the initial axisymmetric phase, 
which lasts several ORPs, the disk cools and contracts, the contraction is small radially but dramatic in the vertical direction. Around 3-4 ORPs the instabilities begin grow to non-linear amplitudes. When this condition is met, the disk undergoes a strong burst of GIs that predominantly manifest themselves in one or two discrete global spiral modes.  The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs. This is then followed by a period of several ORPs where heating temporarily washes out some of the nonaxisymmetry in the disk.  Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.

@@

Figure \ref{fig:Final_Q} shows $Q$, as a function of radius, at $t \sim$ 18 ORP.  Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution.  From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to the growth of nonaxisymmetric structures over this full range.  Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
the azimuthal resolution.  From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.

@@

In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of
all modes present in the disk.  Since the number of modes accessible to the disk is a function of the azimuthal resolution, we expect that higher resolution simulations will manifest more power in higher-order modes 
that behave more locally. We test this by examining the power in various modes to see how they are affected by the choice of grid.

@@

Figure \ref{fig:DensityPlots}, shows midplane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations. Two features are readily apparent; disks run at higher resolution are thicker than lower resolution disks {\bf (anyone want to comment on this?)} and the character of density perturbations differ with resolution.
The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations. Higher angular resolution allows GI power to spread to higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density. The strengths of these modes are given by their global Fourier amplitudes,

@@

Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep{lodato2004,mejia2005,boley2006,cossins2009}.  Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution. While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
the distribution of this power shows marked differences. In particular, the fractional power in lower order modes decreases monotonically with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these modes when $l_{max} =$ 512.

The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution. Thus power is naturally spread from lower order to higher order modes as the available $m$-values increases. Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.

Given the resolution dependence of how power is distributed,  we must examine what resolution is sufficient to properly capture the evolution of these disks. Table \ref{tbl:ams} shows that while the fractional power in lower order modes decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512 
is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution. Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.42 for increasingly higher resolutions. This is displayed graphically in Figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid 
spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
equals unity.  Convergence occurs as the normalized spacing goes to zero. The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.42$ is plotted at a normalized grid spacing  equal to zero. The majority of the modal power in a converged disk resides in higher-order modes. 
%These modes dissipate on shorter wavelengths than low-order modes and hence act more locally.

@@

Low-order spiral modes in $\Delta\rho / \rho$  have longer wavelengths that, if coherent in radius, have longer lever 
arms to produce torques, while high-order modes have relatively short wavelengths which, particularly if they 
lack coherence, produce more localized effects. High-order modes tend to produce pockets of local GI activity, 
which can cancel each other over the whole disk.  The finding that the fractional power in high-order modes increases
with resolution, while the total power stays the same, leads to
 the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.

The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$ can be obtained by integrating the stress tensor $T$ over the surface of the cylinder \citep{lyndenbell1972}, i.e.,
\begin{equation}
{\rm {\bf C}} = \int {\bf r} \times T \cdot {\rm d}S.
\end{equation}
If the stress tensor includes only gravitational stresses, the surface integral can be replaced with the volume integral
\begin{equation}
{\rm \bf C} = \int \rho {\bf r} \times \nabla \Phi {\rm d}V,
\end{equation}
where $\Phi$ is the gravitational potential.  Here we are interested only in the $z$-component of torque,  
\begin{equation}
\label{torque1}
{\rm \bf C}_Z = \int \rho \frac{\partial\Phi} {\partial \phi} {\rm d}V,
\end{equation}
as only this component drives mass and angular momentum transfer. The torque contribution from each mode $m$ can be calculated by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
\begin{equation}
\rho_m = a_{\phi m} \cos(m\phi) + b_{\phi m} \sin(m\phi),
\end{equation}
where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$, and only the gravitational potential produced by the mass distribution given by $\rho_m$ is included in $\Phi$.

Figure 5  displays 
 time-averaged torques summed over a number of low-order modes, $\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, $n = $ 1, 2, 3, 4, and 5, and $\langle {\rm \bf C}_{Ztot} \rangle$,
the total time-averaged gravitational torque arising from modes 1 thru $l_{max}/2$, for four different time-frames,
as a function of radius and azimuthal resolution. 
The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
where the scale is doubled.  
Because the star's position is artificially fixed at the center of the computational grid, 
torques arising from the $m=1$ mode,
typically 5-10\% of the total torque when averaged over the radial range of the figure,
should not be considered realistic.  
Michael \& Durisen (2010) found that freeing the star makes only 10-20\%
differences,  thus the difference between torques arising from the $m=1$ mode with the star fixed are roughly half that found when the star is free to move {\bf (Scott or Dick: please rework this sentence)}.

We note two results of Figure 5: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions. Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque. As resolution increases, non-axisymmetric power is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.

@@

>>>>

We conducted a convergence study of the azimuthal resolution in 3D hydrodynamic simulations of a protostellar disk susceptible to gravitational instabilities (GIs). 
The disk was primarily studied during its asymptotic state, when
heating and cooling are in balance. The cooling was treated as a constant global cooling time.
Four simulations were conducted, identical except for the number of azimuthal grid 
points, $l_{max} = $ 64, 128, 256 and 512, used in the calculations. 
The strengths of the azimuthal modes, gravitational torques, and the effective \citet{shakura1973} $\alpha$ 
arising from gravitational stresses were determined for each resolution.  We examined the resolution-dependence of each quantity
and evaluated both the converged modal amplitude distribution and the effective value of $\alpha$
as $l_{max} \rightarrow \infty$.   Non-axisymmetric amplitude was found in all modes and 
the amplitude summed over all modes was essentially independent of resolution.
The modal distribution of amplitudes, however, was significantly affected by resolution.  
The size of the azimuthal grid, $l_{max}$,
limits the number of modes that can be accurately measured to $l_{max}/2$.  Higher resolution simulations therefore
have more modes available to them.  Non-axisymmetric structure is present in all available modes.  As a consequence,
higher resolution simulations have a larger fraction of their total amplitude in high-order modes.
High-order modes
are more confined to act on local structure than low-order modes, therefore as the resolution increases and low-order amplitudes decrease the total gravitational stresses decreases as well.
As a result, higher resolution simulations experience 
weaker gravitational torques than lower resolution simulations.   The effective  $\alpha$
also depends upon the magnitude of the stresses.  Thus
$\alpha_{\rm{eff}}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
The converged $\alpha_{\rm{eff}}$ is consistent with
predictions from an analytic local theory for thin disks by \citet{gammie2001}, 
but only over many dynamic times when averaged over a substantial volume of the disk.
Results obtained in the $l_{max} = 512$ simulations differ with converged values
by less than 20\%.

@@

Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
modes can become unstable in the linear regime \citep{nelson1998, pickett1998, lodato2004,boley2006,cossins2009}.  Even if the initial growth is dominated by one mode, numerous modes usually appear in the nonlinear regime. As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, referred to as the {\it asymptotic state},
then power will exist at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.  
In the asymptotic state, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
\citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.  Here convergence studies are important because 
higher order modes dissipate energy on shorter wavelengths than lower order modes.  The mix of low and high order modes available to a disk is set by the resolution of the simulation. {\bf Multiple issues with this paragraph. Didn't want to rewrite unilaterally. 1) In the first couple sentences the description of mode growth needs some tweaking, many modes appear in the linear regime, and although we have seen amplitude at all m values I don't know that it's a given. 2) I wouldn't describe the structures formed by GI activity as ``strongly nonlinear turbulence''. 3) In reviewing Gammie 2001 it seems that he is actually referring to a turbulent cascade (a fact I hadn't fully appreciated until now) I think it's pretty clear that the effective alpha we see is, in fact, not due to a turbulent cascade. i.e. alpha is driven by low order coherent modes. We should discuss this at some point SM}

\citet{shakura1973} proposed turbulence as the primary source of effective viscosity in gas disks
and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  The source of the 
turbulence in their $\alpha$-disk formulation is unspecified. Although their original work was directed towards X-ray bright
accretion disks around black holes, this heuristic approach has proven fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling. In the 
$\alpha$-disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $\Sigma$ is the surface density and $H$ represents
the disk scale height. The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
and heating are dominated by local processes}. \citet{lin1987,lin1990} were among the first to suggest
that GI-induced transport could be described within a viscous
framework.  However, the  validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation. 
The locality of dissipation in GI-active disks has not been rigorously tested and remains an open question.

Several studies support the idea that angular momentum transport by GIs in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.  This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance, i.e., disks subject to gravitoturbulence,
 it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.  In particular, 
an $\alpha$-viscosity may be consistent
with transport in GI active disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
in systems with larger $M_d/M_*$ are dominated by lower-order modes that act in a more global sense and thus cannot be well represented by an $\alpha$-approach
\citep{lodato2004, cossins2009, vorobyov2010}.

\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating caused by GIs was balanced by the local cooling rate.  
Assuming a razor thin gravitationally unstable disk where the heating caused by
GIs was balanced by a local cooling rate of $t_{cool}\Omega$.
Gammie found
\begin{equation}
\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1},
\label{eq:gammie}
\end{equation}
where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
The two dimensional index relates
to the three dimensional index, $\Gamma$, by 
$\gamma = (3\Gamma - 1) / (\Gamma + 1)$
for a non-self-gravitating
disk and
$\gamma = 3 - 2/\Gamma $
for a strongly self gravitating disk.   \cite{lodato2004} derived a similar result when considering three dimensional protoplanetary disks.
Motivated by
the results from \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order, \citet{kratter2008} 
proposed $\alpha_{\rm{eff}} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
are functions of $Q$ and $M_d/\left( M_d+M_*\right)$ representing 
the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.

We address these, and similar questions, using a grid-based, finite-difference 3D hydrodynamics code
with self gravity.
For computational grids of a fixed {\it physical} size,
grids with more elements provide finer meshes, and thus a more accurate solution. 
However, increased accuracy is offset by computational penalties.
The number of equations solved at each time step 
increases as the grid size is increased.  At the same time, the Courant condition 
forces shorter time steps.  A mesh convergence study 
provides a common approach to balancing accuracy and computing resources.
In the following sections we follow the evolution of a disk subject to GIs
using four different
resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
may be particularly important, because higher azimuthal resolution 
allows non-axisymmetric amplitude to spread to higher-order modes 
that behave more locally. We test this by examining the amplitude in various 
modes to see how they are affected by the choice of grid.
In the process, 
we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium If this is the case, we investigate how closely these disks approximate the gravitoturbulence of \citet{gammie2001} and examine the distribution of $\alpha_{\rm{eff}}$. To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.  Simulations are performed using four different resolutions in the azimuthal direction.

@@

The initially unstable disk model passed through several phases of evolution \citep[see also][]{pickett2003, mejia2005}.  The initial disk is unstable to the growth of non-axisymmetric structures. During the initial axisymmetric phase, 
which lasts several ORPs, the disk cools and contracts, the contraction is small radially but dramatic in the vertical direction. Around 3-4 ORPs the instabilities begin grow to non-linear amplitudes. When this condition is met, the disk undergoes a strong burst of GIs that predominantly manifest themselves in one or two discrete global spiral modes.  The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs. This is then followed by a period of several ORPs where heating temporarily washes out some of the non-axisymmetry in the disk.  Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.

@@

Figure \ref{fig:Final_Q} shows the Toomre $Q$, as a function of radius, for all four simulations at $t \sim$ 18 ORP.  Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of each of the asymptotic disks, time-averaged from 12 -- 19 ORP and spatially averaged over 10 -- 40 AU.  From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject GIs over this full range.  Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
the azimuthal resolution.  From 27 -- 50 AU, lower resolution simulations display lower values of $Q$. {\bf I suppose 27 AU was chosen by eye? Or was there some other analysis. I vote to toss these last two sentences as I'm not convinced w/o a time averaged quantity. Unless this observation is used in further reasoning in the discussion...SM}

@@

In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of
modes in the disk.  Since the number of modes accessible to the disk depends on the azimuthal resolution, we expect that higher resolution simulations will exhibit non-axisymmetric amplitude in higher-order modes inaccessible at lower resolution, and these higher-order modes will 
behave more locally. We test this by examining the amplitude in various modes to see how they are affected by the choice of grid.

@@

Figure \ref{fig:DensityPlots}, shows mid-plane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations. It is readily apparent that the character of the non-axisymmetric density structures differ with 
resolution. One can clearly see that the lower resolution simulations, i.e. $l_{max} = 64$ and 128, have primarily low-order structures with $m = 2$ and 3 dominating. In contrast, the higher resolution simulations have much more fine structure, the modes are much more washed out, and the low-order modes no longer dominate.
The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations. Higher angular resolution allows non-axisymmetric structures to grow in higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density. The strengths of these modes are given by their global Fourier amplitudes,

@@

Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep{lodato2004,mejia2005,boley2006,cossins2009}.  Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution. We identify $m=2-7$ as the low-order modes that are considered to be global modes, again 
$m=1$ is excluded due to the fixed central star. This determination was made by considering the radial range over 
which the mode is most effective in transporting angular momentum, i.e. from the inner Linblad resonance to the outer 
Linblad resonance, and comparing it to the disk scale height. These quantities are roughly equal for $m=7$, for lower 
order modes the radial range exceeds the scale height.
While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
the distribution of this power shows marked differences. In particular, the fractional power in lower order modes decreases monotonically with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these modes when $l_{max} =$ 512.

The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution. Thus amplitude is naturally spread from lower order to higher order modes as the available $m$-values increase. Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}. {\bf Haven't had a chance to check these refs but I'm reluctant to call what we see a cascade of power. I suggest striking this sentence. SM}

Given how the non-axisymmetric amplitude distribution depends on the azimuthal resolution, we must examine what resolution is sufficient to properly capture the evolution of these disks. Table \ref{tbl:ams} shows that while the fractional power in lower order modes decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512 
is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution. Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.42 for increasingly higher resolutions. This is displayed graphically in figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid 
spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
equals unity.  Convergence occurs as the normalized spacing goes to zero. The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.42$ is plotted at a normalized grid spacing  equal to zero. 
%The majority of the modal power in a converged disk resides in higher-order modes. 
%These modes dissipate on shorter wavelengths than low-order modes and hence act more locally.

@@

Low-order spiral modes in $\Delta\rho / \rho$  have longer wavelengths that, if coherent in radius, have longer lever 
arms to produce torques, while high-order modes have relatively short wavelengths which, particularly if they 
lack coherence can cancel each other out, and produce more localized effects. 
The finding that the fractional power in high-order modes increases
with resolution, while the total power stays the same, leads to
 the expectation that the gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.

The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$ can be obtained by integrating the stress tensor $T$ over the surface of the cylinder \citep{lyndenbell1972}, i.e.,
\begin{equation}
{\rm {\bf C}} = \int {\bf r} \times T \cdot {\rm d}S.
\end{equation}
If the stress tensor includes only gravitational stresses, the surface integral can be replaced with the volume integral
\begin{equation}
{\rm \bf C} = \int \rho {\bf r} \times \nabla \Phi {\rm d}V,
\end{equation}
where $\Phi$ is the gravitational potential.  Here we are interested only in the $z$-component of torque,  
\begin{equation}
\label{torque1}
{\rm \bf C}_Z = \int \rho \frac{\partial\Phi} {\partial \phi} {\rm d}V,
\end{equation}
as only this component drives mass and angular momentum transfer. The torque contribution from each mode $m$ can be calculated by replacing $\rho$ in equation \eqref{torque1} with the density distribution reconstructed from a single Fourier component
\begin{equation}
\rho_m = a_{\phi m} \cos(m\phi) + b_{\phi m} \sin(m\phi),
\end{equation}
where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$, and only the gravitational potential produced by the mass distribution given by $\rho_m$ is included in $\Phi$.

Figure \ref{fig:torquearray} displays 
 time-averaged torques summed over a number of low-order modes, $\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, $n = $ 1, 2, 3, 4, and 5, and $\langle {\rm \bf C}_{Ztot} \rangle$,
the total time-averaged gravitational torque arising from modes 1 thru $l_{max}/2$, for four different time-frames,
as a function of radius and azimuthal resolution. 
The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
where the scale is doubled.  
Because the star's position is artificially fixed at the center of the computational grid, 
torques arising from the $m=1$ mode,
typically 5-10\% of the total torque when averaged over the radial range of the figure,
should not be considered realistic.  
\citet{michael2010} found that freeing the star changes the torques by 10-20\% at most, 
thus the difference between torques arising from the $m=1$ mode with the star fixed and free to move are small compared to the differences found when comparing azimuthal resolution.

We note two results of figure \ref{fig:torquearray}: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions. Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque. As resolution increases, non-axisymmetric amplitude is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.

@@

{\bf I still think the scale in figure 5 is off by a factor of 2 probably due to the fact that you've used a 0.5 $M_\odot$ star. I also vote for a single 12-18 ORP panel. SM}
